### 2021-04-02 input/proposal.tex
wip, on the understanding of properties

%% to better adapt positioning (I expect 2.5 pages
%% intro+related work)
\newpage

@@

Next Section introduces an adaptive scoped broadcast where the state
of each process is dynamic and changes depending all distinct scopes
that exist in the system.

% the cost of
% which actually depends on its operations, and the current scopes that
% exist in the system.

@@

Unfortunately, processes do not share a common global knowledge of the
network state. For process to reach consistent partitioning together,
a process becoming a source must notify every process that is closer
to it than any other source. Epidemic
dissemination~\cite{birman1999bimodal} constitutes an efficient
communication pattern to perform such operation.

\begin{theorem}[\underline{B}est \underline{e}ventual \underline{d}elivery $(BED) \implies CP$]
  Assuming reliable communication links, every process eventually
  delivers, hence forwards, the notification message $\alpha_s$ of its
  best partition $P_s$, eventually leading to consistent partitioning.
\end{theorem}

\begin{proof} When a process $s_1$ becomes a source, it belongs to
its own partition, for there exists no better partition than its own:
$\forall p \in V: |\Pi_{s_1 s_1}| < |\Pi_{s_1 p}|$. It broadcasts such
message to its neighbors. Since communication channels are reliable,
neighboring processes eventually receive such notification. In
particular, whatever the order of received messages, every process
$p_1$ such that there exists no better partition $s_2$ than the
received one delivers and forwards it: $\forall s_2 \in S: |\Pi_{p
  s_1}| < |\Pi_{p s_2}|$. The notification emanating from $s_1$
transitively reaches all such processes through shortest paths. Since
there exists only one best value per process that can never be
retracted, processes eventually reach consistent partitioning.
\end{proof}

%% \input{input/figdel.tex} %% figure positioning

%% \begin{definition}[\label{def:transitivedelivery}Transitive delivery]
%%   Transitive delivery ensures the system eventually converges to
%%   optimal partitioning:
%%   \begin{inparaenum}[(i)]
%%   \item (\textbf{termination}) processes do not need to propagate
%%     operations they do not deliver: $\forall \langle i, j \rangle \in
%%     E, \neg d_i(\alpha_{s}^w)\implies \neg
%%     d_j(\alpha_{s}^{w+w_{ij}})$;
%%   \item (\textbf{propagation}) operations delivered by a process may
%%     also benefit neighbor processes: $\forall \langle i, j \rangle \in
%%     E, d_i(\alpha_{s}^{w_1}) \implies
%%     d_j(\alpha_{s}^{w_1+w_{ij}}) \vee
%%     d_j(\alpha_{*}^{w_2 < w_1 + w_{ij}})$.
%%   \end{inparaenum}
%% \end{definition}

@@

Algorithm~\ref{algo:add} shows the instructions that implement
eventually consistent partitioning when
\begin{inparaenum}[(i)]
\item processes only have local knowledge about their direct
  neighborhoods,
\item weights are scalar values,
\item and processes only add new partitions to the system.
\end{inparaenum}
At any time, a process running this protocol can add a logical
partition to its distributed system.  Figure~\ref{fig:add} illustrates
its behavior on a system comprising 4 processes $a$, $b$, $c$, and
$d$. Two processes $a$ and $c$ concurrently create a partition by
S-broadcasting their respective \underline{a}dd message: $\alpha_a^0$
and $\alpha_c^0$. They initialize their own state with the lowest
possible bound $0$ (see Line~\ref{line:lowestbound}), and send a
message to each of their neighbors by accumulating the corresponding
edge weight (see Line~\ref{line:accumulator}). In
Figure~\ref{fig:addC}, Process $b$ receives $\alpha_{c}^{1}$. Since it
improves its own partition distance, it keeps it and forwards it to
its neighbors. Process $b$ discards $\alpha_{b}^{1.5}$, for it does
not improve its partition distance. Processes $c$ and $d$ will never
hear of Process $a$'s partition.  In Figure~\ref{fig:addD}, processes
discard last transiting messages. The system converged to the
consistent partitioning. However, since processes operate using local
knowledge only, generated traffic may be sub-optimal, for processes
may receive, deliver, and forward messages emanating from a same
source simply because it did not come from the shortest path yet.

@@

\subsection{Dynamic consistent partitioning}
At any time, a process can become a source, hence adding a new
partition to the system. This partition eventually includes all
processes that are closer from this new source than any other else. We
described such protocol in Section~\ref{sec:scoped}. Processes
naturally converge towards their respective best partition by only
piggybacking a monotonically increasing distance in forwarded
messages. % \TODO{Traffic of each partition is contained to the
%  partition.}

@@

\begin{theorem}[\label{theo:dcp}Dynamic consistent partitioning]
  Eventual consistent partitioning in presence of dynamic sources
  requires
  \begin{inparaenum}[(i)]
  \item eventual purging of stale notifications, and
  \item eventual delivery of bests partitions.
  \end{inparaenum}
\end{theorem}

>>>>

%% to better adapt positioning (I expect 2.5 pages
%% intro+related work)

\vfill \TODO{spacing for intro and related work: 2.5 pages}
\newpage

@@

Next Section introduces an adaptive scoped broadcast where the state
of each process is dynamic and changes depending on all distinct
scopes that exist in the system.

@@

Unfortunately, processes do not share a common global knowledge of the
network state. For processes to reach consistent partitioning
together, a process becoming a source must notify every process that
is closer to it than any other source. Epidemic
dissemination~\cite{birman1999bimodal} constitutes an efficient
communication pattern to perform such operation.

\begin{theorem}[\underline{B}est \underline{e}ventual \underline{f}orwarding $(BEF) \iff CP$]
  Assuming reliable communication links where a correct process $q$
  eventually receives the message $m$ \underline{s}ent to it by a
  process $p$ ($s_{pq}(m) \implies r_{q}(m)$); a static system ($G^t =
  G^{t+1}$); assuming local knowledge only, processes eventually reach
  consistent partitioning if and only if each process eventually
  forwards its best known partition.
\end{theorem}

\TODO{If forwarding defined as send to all neighbors, then it also
  works in dynamic systems where processes join and edges are added,
  for when a new neighbor appears, the processes must forward their
  best.}

\begin{proof}
  \begin{asparadesc}
  \item [$BEF \implies CP$:] When a process $s_1$ becomes a source, it
    belongs to its own partition, for there exists no better partition
    than its own: $\forall p \in V: |\Pi_{s_1 s_1}| < |\Pi_{s_1
      p}|$. It broadcasts such message to its neighbors. Since
    communication channels are reliable, neighboring processes
    eventually receive such notification. In particular, whatever the
    order of received messages, every process $p_1$ such that there
    exists no better partition $s_2$ than the received one delivers
    and forwards it: $\forall s_2 \in S: |\Pi_{p s_1}| < |\Pi_{p
      s_2}|$. The notification emanating from $s_1$ transitively
    reaches all such processes through shortest paths. Since there
    exists only one best value per process that can never be
    retracted, processes eventually reach consistent
    partitioning. \TODO{Rework this with $s_{pq}$ and $r_{pq}$.}
  \item [$CP \implies BEF$:] \TODO{By contradiction, if there is a
    process $p$ that does not forward, then another process $q$ may
    never gets its best because it needs the weight $W_{pq}$.}
  \end{asparadesc}
\end{proof}

@@

Algorithm~\ref{algo:add} shows the instructions that implement such
consistent partitioning when
\begin{inparaenum}[(i)]
\item weights are scalar values,
\item processes only add new partitions to the system,
\item and processes never crash nor leave the system.
\end{inparaenum}
 Figure~\ref{fig:add} illustrates its behavior on a system comprising
 4 processes $a$, $b$, $c$, and $d$. Two processes $a$ and $c$
 concurrently create a partition by S-broadcasting their respective
 \underline{a}dd message: $\alpha_a^0$ and $\alpha_c^0$. They
 initialize their own state with the lowest possible bound $0$ (see
 Line~\ref{line:lowestbound}), and send a message to each of their
 neighbors by accumulating the corresponding edge weight (see
 Line~\ref{line:accumulator}). In Figure~\ref{fig:addC}, Process $b$
 receives $\alpha_{c}^{1}$. Since it improves its own partition
 distance, it keeps it and forwards it to its neighbors. Process $b$
 discards $\alpha_{b}^{1.5}$, for it does not improve its partition
 distance. Processes $c$ and $d$ will never hear of Process $a$'s
 partition.  In Figure~\ref{fig:addD}, processes discard last
 transiting messages. The system converged to the consistent
 partitioning. However, since processes operate using local knowledge
 only, generated traffic may be sub-optimal, for processes may
 receive, deliver, and forward messages emanating from a same source
 simply because it did not come from the shortest path yet.

@@

\subsection{Dynamic consistent partitioning}
At any time, a process can become a source, hence adding a new
partition to the system. This partition eventually includes all
processes that are closer from this new source than any other
else. Processes naturally converge towards their respective best
partition by only piggybacking a monotonically increasing distance in
forwarded messages.

@@

\begin{corollary}[\label{theo:dcp}Dynamic consistent partitioning]
  Eventual consistent partitioning in presence of dynamic sources
  requires
  \begin{inparaenum}[(i)]
  \item eventual purging of stale notifications, and
  \item eventual delivery of bests partitions.
  \end{inparaenum}
\end{corollary}
